---
title: "Final Project"
author: "Rachel Feingold"
date: "5/1/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Wildfires 

## Packages Used 
These are the pachages required by my code to run
```{r}
require(foreign)
require(ff)
require(dplyr)
require(lubridate)
require(ggplot2)
require(ggmap)
require(dplyr)
```

## Loading Data 

I decide to start with the first year in the Google Drive, 2007. 

```{r}
wild07 <- read.dbf("~/MA415/Final Project/2007/wildlands.dbf")
codes <-  read.dbf("~/MA415/Final Project/2007/codelookup.dbf")
```

Cleaning the 2007 Data

I used two primary functions: table(factor()) and length(x[is.na(x)]) to pick and choose which variables I would keep to observe.
```{r}
# keeping the relevant columns from wild07
length(wild07$HUM_FACT1[is.na(wild07$HUM_FACT1)])

table(factor(wild07$FIRE_CAUSE))
# All NA's for human factors 7 and 8, most variation for human factor 1
table(factor(wild07$HUM_FACT1))
# table(factor(wild07$HUM_FACT2))
# table(factor(wild07$HUM_FACT3))
# table(factor(wild07$HUM_FACT4))
# table(factor(wild07$HUM_FACT5))
# table(factor(wild07$HUM_FACT6))

table(factor(wild07$FACT_IGN1))
table(factor(wild07$FACT_IGN2))

# table(factor(wild07$SUPP_FACT1))
# table(factor(wild07$SUPP_FACT2))
# table(factor(wild07$SUPP_FACT3))

length(wild07$BLDG_INV[is.na(wild07$BLDG_INV)])
mean(wild07$BLDG_INV,na.rm = T) # 0.074
mean(wild07$BLDG_THR,na.rm = T) # 0.3077
mean(wild07$ACRES_BURN,na.rm = T) # 466.4983

# table(factor(wild07$CROP_BURN1))
# table(factor(wild07$CROP_BURN2))
# table(factor(wild07$CROP_BURN3))

length(wild07$UNDET_BURN[is.na(wild07$UNDET_BURN)])
mean(wild07$UNDET_BURN,na.rm = T) # 32.96
mean(wild07$TAX_BURN,na.rm = T) # 10.12

table(factor(wild07$PROP_MANAG))

mean(wild07$ELEVATION, na.rm = T) # 157.87 
mean(wild07$FLAME_LGTH, na.rm = T) # 0.439
mean(wild07$SPREAD_RAT, na.rm = T) # 0.562


wild07 <-  select(wild07,STATE, FDID, INC_DATE, INC_NO, LATITUDE, LONGITUDE, AREA_TYPE, FIRE_CAUSE, HUM_FACT1, FACT_IGN1, FACT_IGN2,
                   ACRES_BURN, UNDET_BURN, PERSON_FIR, GENDER, AGE, ACTIVITY_W, ELEVATION, FLAME_LGTH, SPREAD_RAT)

```
Then, I transformed the character variables that had coding for different factors using the codes.dbf file.
```{r}

# Want AREA-TYPE to have a real, written desciption, instead of number for each type of area.
a_types  <- codes[(codes$FIELDID == "AREA_TYPE"),]
a_types <- rename(a_types, AREA_TYPE = CODE_VALUE)
wild07 <- left_join(wild07, a_types, by = "AREA_TYPE")
wild07 <- select(wild07, -AREA_TYPE, -FIELDID)
wild07 <- rename(wild07, AREA_TYPE = CODE_DESCR)  
rm(a_types)

# Do the same for Fire Cause
fc <- codes[(codes$FIELDID == "FIRE_CAUSE"),]
fc <- rename(fc, FIRE_CAUSE = CODE_VALUE)
wild07 <- left_join(wild07, fc, by = "FIRE_CAUSE")
wild07 <- select(wild07, -FIRE_CAUSE, -FIELDID)
wild07 <- rename(wild07, FIRE_CAUSE = CODE_DESCR)  
rm(fc)

# Do the same for Human Factors 1 Contributing to Ignition
hf1 <- codes[(codes$FIELDID == "HUM_FAC_1"),]
hf1 <- rename(hf1, HUM_FACT1 = CODE_VALUE)
wild07 <- left_join(wild07, hf1, by = "HUM_FACT1")
wild07 <- select(wild07, -HUM_FACT1, -FIELDID)
wild07 <- rename(wild07, HUM_FACT1 = CODE_DESCR)  
rm(hf1)

# Do the same for factors Contributing to Ignition (1 and 2)

if1 <- codes[(codes$FIELDID == "FACT_IGN_1"),]
if1 <- rename(if1, FACT_IGN1 = CODE_VALUE)
wild07 <- left_join(wild07, if1, by = "FACT_IGN1")
wild07 <- select(wild07, -FACT_IGN1, -FIELDID)
wild07 <- rename(wild07, FACT_IGN1 = CODE_DESCR)  
rm(if1)

if2 <- codes[(codes$FIELDID == "FACT_IGN_2"),]
if2 <- rename(if2, FACT_IGN2 = CODE_VALUE)
wild07 <- left_join(wild07, if2, by = "FACT_IGN2")
wild07 <- select(wild07, -FACT_IGN2, -FIELDID)
wild07 <- rename(wild07, FACT_IGN2 = CODE_DESCR)  
rm(if2)

# Same procedure for demographic data of people who ignited the fires

ppl <- codes[(codes$FIELDID == "PERSON_FIR"),]
ppl <- rename(ppl, PERSON_FIR = CODE_VALUE)
wild07 <- left_join(wild07, ppl, by = "PERSON_FIR")
wild07 <- select(wild07, -PERSON_FIR, -FIELDID)
wild07 <- rename(wild07, PRSN_RESPONS = CODE_DESCR)  
rm(ppl)

gen <- codes[(codes$FIELDID == "GENDER"),]
gen <- rename(gen, GENDER = CODE_VALUE)
wild07 <- left_join(wild07, gen, by = "GENDER")
wild07 <- select(wild07, -GENDER, -FIELDID)
wild07 <- rename(wild07, GENDER = CODE_DESCR)  
rm(gen)

# Do for Activity performed by the person involved

act <- codes[(codes$FIELDID == "ACTIVITY_W"),]
act <- rename(act, ACTIVITY_W = CODE_VALUE)
wild07 <- left_join(wild07, act, by = "ACTIVITY_W")
wild07 <- select(wild07, -ACTIVITY_W, -FIELDID)
wild07 <- rename(wild07, IGN_ACTIVITY = CODE_DESCR)  
rm(act)

# Wanted to change the format of the Incident Dates and make them more sensible to read. 
wild07$INC_DATE <- mdy(wild07$INC_DATE)
```
Then, I wanted to reorder columns into more sensible train of thought. First, I have variables that describe the incident, like incident number, incident ID, date of inicident, and information about the location of the event. Then, I have variables describing the severity and extent of the incident. After, I have variables concerning the factors contributing to cause of ignition. If a human started the incident, then I ordered variables that described the demographics of the person.
```{r}
wild07 <- wild07[,c("INC_NO", "FDID", "INC_DATE", "STATE", "LATITUDE","LONGITUDE", "AREA_TYPE", "ACRES_BURN", "UNDET_BURN", "FLAME_LGTH","SPREAD_RAT", "ELEVATION", "FIRE_CAUSE", "FACT_IGN1", "FACT_IGN2", "HUM_FACT1", "IGN_ACTIVITY", "PRSN_RESPONS", "GENDER", "AGE")]

```

## Similar Procedure to Load and Clean Data from the year 2015

I needed to use read table because the data was in txt files.
```{r}
wild15 <- read.table("~/MA415/Final Project/2015/wildlands.txt", header = T, sep = "^")
codes15 <- read.table("~/MA415/Final Project/2015/codelookup.txt", header = T, sep = "^")
```

##Cleaning Data

```{r}
#Selecting variables of interest
wild15 <-  select(wild15,STATE, FDID, INC_DATE, INC_NO, LATITUDE, LONGITUDE, AREA_TYPE, FIRE_CAUSE, HUM_FACT1, FACT_IGN1, FACT_IGN2,
                   ACRES_BURN, UNDET_BURN, PERSON_FIR, GENDER, AGE, ACTIVITY_W, ELEVATION, FLAME_LGTH, SPREAD_RAT)

# Quick Stats 
mean(wild15$ACRES_BURN,na.rm = T) # 274.13
mean(wild15$UNDET_BURN,na.rm = T) # 34.66
mean(wild15$ELEVATION, na.rm = T) # 292.6005 
mean(wild15$FLAME_LGTH, na.rm = T) # 0.49
mean(wild15$SPREAD_RAT, na.rm = T) # 0.678

# Transforming the character variables using the codes.dbf file.

# Want AREA-TYPE to have a real, written desciption, instead of number for each type of area. 
# But area_type reads as numeric when you read the file in rith read.table(), need to change it to a factor first
wild15$AREA_TYPE <- as.factor(wild15$AREA_TYPE)
a_types  <- codes15[(codes15$fieldid == "AREA_TYPE"),]
a_types <- rename(a_types, AREA_TYPE = code_value)
wild15 <- left_join(wild15, a_types, by = "AREA_TYPE")
wild15 <- select(wild15, -AREA_TYPE, -fieldid)
wild15 <- rename(wild15, AREA_TYPE = code_descr)  
rm(a_types)

# Do the same for Fire Cause
wild15$FIRE_CAUSE <- as.factor(wild15$FIRE_CAUSE)
fc  <- codes15[(codes15$fieldid == "FIRE_CAUSE"),]
fc <- rename(fc, FIRE_CAUSE = code_value)
wild15 <- left_join(wild15, fc, by = "FIRE_CAUSE")
wild15 <- select(wild15, -FIRE_CAUSE, -fieldid)
wild15 <- rename(wild15, FIRE_CAUSE = code_descr)  
rm(fc)

# Do the same for Human Factors 1 Contributing to Ignition
wild15$HUM_FACT1 <- as.factor(wild15$HUM_FACT1)
hf1  <- codes15[(codes15$fieldid == "HUM_FACT1"),]
hf1  <- rename(hf1, HUM_FACT1 = code_value)
wild15 <- left_join(wild15, hf1, by = "HUM_FACT1")
wild15 <- select(wild15, -HUM_FACT1, -fieldid)
wild15 <- rename(wild15, HUM_FACT1 = code_descr)  
rm(hf1)

# Do the same for factors Contributing to Ignition (1 and 2)
wild15$FACT_IGN1 <- as.factor(wild15$FACT_IGN1)
ign  <- codes15[(codes15$fieldid == "FACT_IGN_1"),]
ign  <- rename(ign, FACT_IGN1 = code_value)
wild15 <- left_join(wild15, ign, by = "FACT_IGN1")
wild15 <- select(wild15, -FACT_IGN1, -fieldid)
wild15 <- rename(wild15, FACT_IGN1 = code_descr)  
rm(ign)

wild15$FACT_IGN2 <- as.factor(wild15$FACT_IGN2)
ign  <- codes15[(codes15$fieldid == "FACT_IGN_2"),]
ign  <- rename(ign, FACT_IGN2 = code_value)
wild15 <- left_join(wild15, ign, by = "FACT_IGN2")
wild15 <- select(wild15, -FACT_IGN2, -fieldid)
wild15 <- rename(wild15, FACT_IGN2 = code_descr)  
rm(ign)

# Same procedure for demographic data of people who ignited the fires

wild15$PERSON_FIR <- as.factor(wild15$PERSON_FIR)
ppl  <- codes15[(codes15$fieldid == "PERSON_FIR"),]
ppl  <- rename(ppl, PERSON_FIR = code_value)
wild15 <- left_join(wild15, ppl, by = "PERSON_FIR")
wild15 <- select(wild15, -PERSON_FIR, -fieldid)
wild15 <- rename(wild15, PRSN_RESPONS = code_descr)  
rm(ppl)

wild15$GENDER <- as.factor(wild15$GENDER)
ppl  <- codes15[(codes15$fieldid == "GENDER"),]
ppl  <- rename(ppl, GENDER = code_value)
wild15 <- left_join(wild15, ppl, by = "GENDER")
wild15 <- select(wild15, -GENDER, -fieldid)
wild15 <- rename(wild15, GENDER = code_descr)  
rm(ppl)

# Do for Activity performed by the person involved

wild15$ACTIVITY_W <- as.factor(wild15$ACTIVITY_W)
ppl  <- codes15[(codes15$fieldid == "ACTIVITY_W"),]
ppl  <- rename(ppl, ACTIVITY_W = code_value)
wild15 <- left_join(wild15, ppl, by = "ACTIVITY_W")
wild15 <- select(wild15, -ACTIVITY_W, -fieldid)
wild15 <- rename(wild15, IGN_ACTIVITY = code_descr)  
rm(ppl)

# Wanted to change the format of the Incident Dates and make them more sensible to read. 
wild15$INC_DATE <- mdy(wild15$INC_DATE)

# Change the order of the variables with the same thinking in mind. 

wild15 <- wild15[, c("INC_NO", "FDID", "INC_DATE", "STATE", "LATITUDE", "LONGITUDE", "AREA_TYPE", "ACRES_BURN", "UNDET_BURN", "FLAME_LGTH","SPREAD_RAT", "ELEVATION", "FIRE_CAUSE", "FACT_IGN1", "FACT_IGN2", "HUM_FACT1", "IGN_ACTIVITY", "PRSN_RESPONS", "GENDER", "AGE")]
```
##  Data Plots
I wanted to look at total acres burned per state and see how this changes over time. Do some states have more fire in 2007 than in 2015 and why?
```{r}
dat1 <- select(wild07, STATE, AREA_TYPE, ACRES_BURN)
dat1 <- filter(dat1, ACRES_BURN > 1)
ggplot(dat1) + geom_density(aes(x = ACRES_BURN)) + ggtitle("ggplot of ACRES BURNED by STATE")


dat2 
```
## Data Tables
I wanted to look at the count of different causes of wildfire for each state.

```{r}

```

## Playing with Maps 

I'd like to map different causes of fires over the map of the US, to see if one part of the country has more fires than another.
```{r}
map1 <- get_map(location = "united states", zoom = 4, maptype = "terrain", source = "google", color = "color")
# This is just a map of the US
#ggmap(map1)

# Mapping the wildfire incidents, showing the different causes related to the fire, by colored points.
data <- select(wild07, LATITUDE, LONGITUDE, FIRE_CAUSE)
# If there were NA values for these character variables, the codes set their value as the name of the variable in all caps; we want to filter these out. 
data <- filter(data, FIRE_CAUSE != "WILDLAND FIRE CAUSE")
# add points
m1 <- ggmap(map1) + geom_point(aes(x = LONGITUDE, y = LATITUDE, color = FIRE_CAUSE), data = data, alpha = .5, na.rm = T) 
rm(data)
# Mapping the wildfire incidents, showing the different human factors related to fire ignition, by colored points.
data1 <- select(wild07, LATITUDE, LONGITUDE, HUM_FACT1)
data1 <- filter(data1, HUM_FACT1 != "HUMAN FACTORS CONTRIBUTING TO IGNITION 1")
data1 <- filter(data1, HUM_FACT1 != "None")
m2 <- ggmap(map1) + geom_point(aes(x = LONGITUDE, y = LATITUDE, color = HUM_FACT1), data = data1, alpha = .5, na.rm = T) 
rm(data1)

```
One interesting observation is that in the southern states, most of the human factors were a result of an unattended or unsupervised person.
Next, I'd like to look at total acres burned vs cause of fire.
```{r}

data2 <- select(wild07, LATITUDE, LONGITUDE, ACRES_BURN, HUM_FACT1)
# If there were NA values for these character variables, the codes set their value as the name of the vriable in all caps; we want to filter these out. We also filtered out the data where the human factore was NONE. 
data2 <- filter(data2, HUM_FACT1 != "HUMAN FACTORS CONTRIBUTING TO IGNITION 1")
data2 <- filter(data2, HUM_FACT1 != "None")
data2 <- filter(data2, ACRES_BURN != 0)
# add points
ggmap(map1) + geom_point(aes(x = LONGITUDE, y = LATITUDE, color = HUM_FACT1), data = data1, alpha = .5, na.rm = T) 

```
